---
title: "Respiratory Report"
output: 
  flexdashboard::flex_dashboard:
     theme: cosmo
     orientation: rows
     vertical_layout: fill
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(readr)
library(dplyr)
library(dbplyr)
library(lubridate)
library(readxl)
library(ggplot2)
library(plotly)
library(shiny)
library(magrittr)
library(stringr)
library(flexdashboard)
library(kableExtra)
library(scales)
library(tidyverse)
```

```{r data_loading, echo = F, message = F}
#==============================================================#
# Load Data
#==============================================================#

# Load Respiratory Data
respiratory <- readRDS("N:/Data/rawData/activityData.rda") %>%
  rename_with(snakecase::to_snake_case)

#Load Montelukast data
montelukast <- read.csv(file = 'N:/Data/Montelukast.csv') %>%
  rename_with(snakecase::to_snake_case)

# Load BNF chapter spend data
bnf_spend <- read.csv(file = 'N:/Data/Spend_by_BNFchapter.csv') %>%
  rename_with(snakecase::to_snake_case)

#load a list of active GPs during the reporting period and map their localities
source("N:/03-Reference/R Functions/loadPharmacyFunctions.R")

#Load practice populations
practiceLookup <- read_excel("N:\\03-Reference\\Lookup Information\\practice_information.xlsx") %>%
  select(PracticeCode, practice_population, locality, cluster)

# Load inhaler types
inhaler_types <- read_excel("N:\\03-Reference\\Lookup Information\\DPI vs MDI inhalers.xlsx")
```

```{r data_processing, echo = F, message = F}
#==============================================================#
# Process Data
#==============================================================#

# Join prescribing data to practiceLookup to match on locality/cluster
respiratory %<>% left_join(practiceLookup,
                           by = c("presc_location_code" = "PracticeCode"))

# Create objects for dates
current_month <- max(respiratory$date)
latest_date <- format.Date(current_month, "%B %Y")
full_back12Mth <- format(as.Date(current_month) %m-% months(11), "%Y-%m-%d")
full_back12Mth_pretty <- format(as.Date(current_month) %m-% months(11), "%B %Y")
full_back3Mth <- format(as.Date(current_month) %m-% months(2), "%Y-%m-%d")
full_back3Mth_pretty <- format(as.Date(current_month) %m-% months(2), "%B %Y")
#full_current_month <- format(as.Date(current_month), "%B %Y")
latest_full_quarter <- respiratory %>% group_by(fin_quarter) %>%
  summarise(months_in_quarter = n_distinct(date)) %>%
  filter(months_in_quarter == 3) %>%
  filter(fin_quarter == max(fin_quarter)) %>% select(fin_quarter)
latest_incomplete_quarter <- respiratory %>% group_by(fin_quarter) %>%
  summarise(months_in_quarter = n_distinct(date)) %>%
  filter(months_in_quarter < 3) %>% select(fin_quarter)

# Change data classes for bnf spend
bnf_spend %<>%
  mutate(pd_paid_gic_excl_bb = as.numeric(gsub(",","", bnf_spend$pd_paid_gic_excl_bb)),
         cost = paste(format(round(pd_paid_gic_excl_bb / 1e6,1), trim = T), "M"),
         label = paste(pi_bnf_chapter_description, cost, sep = "\n"))
```

```{r data_summaries, echo = F, message = F}
#==============================================================#
# Summarise Data - High Level
#==============================================================#

# Number of respiratory patients in last 12 months
total_respiratory_patients <- respiratory %>% filter(date >= full_back3Mth) %>%
  summarise(patients = n_distinct(pat_upi_c))

#------------------------------- Spend on Inhalers----------------------------#

# Total spend on inhalers and number of inhalers prescribed in last 12 months in Lothian
inhaler_spend <- respiratory %>% filter(date >= full_back12Mth) %>%
  summarise(Cost = priceR::format_currency(sum(pd_paid_gic_excl_bb),"£"),
            `Number of Items` = formattable::comma(round(sum(number_of_paid_items)),0),
            `Quantity` = formattable::comma(round(sum(paid_quantity)),0))

# Spend on inhalers by GP cluster - most recent 3 months
respiratory_spend_bycluster <- respiratory %>% filter(date >= full_back3Mth, locality != "No HSCP Mapping Found", !is.na(locality)) %>%
  group_by(cluster) %>%
  summarise(respiratory_patients = n_distinct(pat_upi_c),
            cost = sum(pd_paid_gic_excl_bb),
            cost_per_treated_patient = cost / respiratory_patients)

# Items and inhaler cost per treated patient by GP practice
respiratory_bypractice <- respiratory %>% filter(date >= full_back3Mth, locality != "No HSCP Mapping Found", !is.na(locality)) %>%
  group_by(cluster, presc_location_name) %>%
  summarise(respiratory_patients = n_distinct(pat_upi_c),
            cost = sum(pd_paid_gic_excl_bb),
            cost_per_treated_patient = cost / respiratory_patients,
            items = sum(number_of_paid_items),
            items_per_treated_patient = items / respiratory_patients)

#------------------------------------ SABA -----------------------------------#

# Table summarising total patients, patients who have received a SABA,
# patients who only received a SABA, patients who received an ICS inhaler, 
#patients who only received an ICS
single_saba_use <- respiratory %>% filter(date >= full_back12Mth) %>%
  summarise(health_board = "NHS Lothian",
            total_patients = n_distinct(pat_upi_c),
            all_SABA_patients = n_distinct(pat_upi_c[respiratory$inhaler_class == "SABA"]),
            only_SABA_patients = n_distinct(setdiff(pat_upi_c[respiratory$inhaler_class == "SABA"],
                                                    pat_upi_c[respiratory$inhaler_class != "SABA"])),
            all_ICS_patients = n_distinct(pat_upi_c[respiratory$inhaler_class %in% c("ICS", "ICS/LABA")]),
            only_ICS_patients = n_distinct(setdiff(pat_upi_c[respiratory$inhaler_class %in% c("ICS", "ICS/LABA")], 
                                                   pat_upi_c[respiratory$inhaler_class %in% c("LABA", "LAMA", "LAMA/LABA Mast Cell Stabaliser", "SABA", "SAMA", "Triple Therapy")]))
  )
# Broken down by cluster and practice
single_saba_by_practice <- respiratory %>% filter(date >= full_back12Mth, !is.na(cluster), cluster != "No HSCP Mapping Found") %>%
  group_by(cluster, presc_location_code, presc_location_name) %>%
  summarise(#health_board = "NHS Lothian",
    total_patients = n_distinct(pat_upi_c),
    all_SABA_patients = n_distinct(pat_upi_c[respiratory$inhaler_class == "SABA"]),
    only_SABA_patients = n_distinct(setdiff(pat_upi_c[respiratory$inhaler_class == "SABA"],
                                            pat_upi_c[respiratory$inhaler_class != "SABA"])),
    all_ICS_patients = n_distinct(pat_upi_c[respiratory$inhaler_class %in% c("ICS", "ICS/LABA")]),
    only_ICS_patients = n_distinct(setdiff(pat_upi_c[respiratory$inhaler_class %in% c("ICS", "ICS/LABA")], 
                                           pat_upi_c[respiratory$inhaler_class %in% c("LABA", "LAMA", "LAMA/LABA Mast Cell Stabaliser", "SABA", "SAMA", "Triple Therapy")]))
  ) %>%
  mutate(only_SABA_patients_percentage = round((only_SABA_patients/all_SABA_patients * 100),2)) %>%
  arrange(desc(only_SABA_patients_percentage), desc(only_SABA_patients))

# Percentage of patients prescribed > 12 SABA and > 6 SABA per annum
SABAoveruse <- respiratory %>% filter(date >= full_back12Mth, str_detect(inhaler_class, "SABA"), locality != "No HSCP Mapping Found", !is.na(locality)) %>%
  group_by(locality, presc_location_name, pat_upi_c) %>%
  summarise(quantity = sum(adjusted_quantity)) %>%
  mutate(overuse12_flag = ifelse(quantity > 12, 1, 0),
         overuse6_flag = ifelse(quantity > 6, 1, 0)) %>%
  ungroup() %>%
  group_by(locality, presc_location_name) %>%
  summarise(patients = n_distinct(pat_upi_c),
            overuse12SABA = sum(overuse12_flag),
            overuse6SABA = sum(overuse6_flag)) %>%
  mutate(percent12 = round(overuse12SABA / patients * 100,2),
         percent6 = round(overuse6SABA / patients * 100,2)) %>%
  select(-patients) %>%
  arrange(desc(percent12), desc(percent6))

# Formatted table of SABA overuse
SABAoveruse_formatted <- SABAoveruse %>%
  rename(`Number of Patients > 12 SABA in last 12 months` = overuse12SABA,
         `Number of Patients > 6 SABA in last 12 months` = overuse6SABA,
         `Percent of Patients > 12 SABA in last 12 months` = percent12,
         `Percent of Patients > 6 SABA in last 12 months` = percent6)

# Grouping SABA overuse by year
SABAoveruse_by_year <- respiratory %>% filter(str_detect(inhaler_class, "SABA"), locality != "No HSCP Mapping Found", !is.na(locality)) %>%
  mutate(back12month = format(as.Date(date) %m-% months(11), "%Y-%m-%d")) %>%
  group_by(fin_year, locality, presc_location_name, pat_upi_c) %>%
  summarise(quantity = sum(adjusted_quantity)) %>%
  mutate(overuse12_flag = ifelse(quantity > 12, 1, 0),
         overuse6_flag = ifelse(quantity > 6, 1, 0)) %>%
  ungroup() %>%
  group_by(fin_year, locality, presc_location_name) %>%
  summarise(patients = n_distinct(pat_upi_c),
            overuse12SABA = sum(overuse12_flag),
            overuse6SABA = sum(overuse6_flag)) %>%
  mutate(percent12 = overuse12SABA / patients * 100,
         percent6 = overuse6SABA / patients * 100)

# Number of Patients on > 200 doses of SABA MDI (200 doses = 1 inhaler), broken down by practice
SABAdoses <- respiratory %>% left_join(inhaler_types, by = "pi_bnf_item_description") %>% 
  filter(str_detect(inhaler_class, "SABA"), locality != "No HSCP Mapping Found", !is.na(locality), 
         date > full_back3Mth, 
         saba_flag == "SABA",
         category == "MDI") %>%
  group_by(locality, presc_location_name) %>%
  summarise(num_SABA_MDI_patients = n_distinct(pat_upi_c),
            num_SABA_MDI_patients_rx_over200doses = n_distinct(pat_upi_c[adjusted_quantity >= 2]),
            num_SABA_MDI_patients_rx_over300doses = n_distinct(pat_upi_c[adjusted_quantity >= 3]),
            num_SABA_MDI_patients_rx_over400doses = n_distinct(pat_upi_c[adjusted_quantity >= 4])) %>%
  mutate(percent = round(num_SABA_MDI_patients_rx_over200doses / num_SABA_MDI_patients, 2),
         `Percentage of SABA MDI Prescriptions of more than 2 inhalers` = formattable::percent(percent)) %>%
  select(-percent) %>%
  arrange(desc(num_SABA_MDI_patients_rx_over200doses)) %>%
  rename(`Number of Patients with a SABA MDI Prescription in last 3 months` = num_SABA_MDI_patients,
         `Number of Patients with a Prescription of more than 2 SABA MDI Inhalers in last 3 months` = num_SABA_MDI_patients_rx_over200doses,
         `Number of Patients with a Prescription of more than 3 SABA MDI Inhalers in last 3 months` = num_SABA_MDI_patients_rx_over300doses,
         `Number of Patients with a Prescription of more than 4 SABA MDI Inhalers in last 3 months` = num_SABA_MDI_patients_rx_over400doses)

#------------------ Percentage of ICS classified as high strength ----------------#

# Summary of ICS high strength prescribing by practice since Jan 2018
ics_prescribed <- respiratory %>% filter(date >= "2018-01-01", inhaler_class %in% c("ICS", "ICS/LABA", "Triple Therapy"), locality != "No HSCP Mapping Found", !is.na(locality)) %>%
  group_by(fin_quarter, locality, presc_location_code, presc_location_name, inhaler_class, hscs_class) %>%
  summarise(quantity = sum(adjusted_quantity)) %>%
  mutate(highstrength_flag = ifelse(str_detect(hscs_class, "High"), "High", "Other")) %>%
  ungroup() %>%
  group_by(fin_quarter, locality, presc_location_code, presc_location_name) %>%
  summarise(highstrengthquantity = sum(quantity[highstrength_flag == "High"]),
            total_quantity = sum(quantity),
            percent_highstrength = highstrengthquantity / total_quantity * 100) %>%
  ungroup()

# Formatted table of ICS prescribing by practice - last full Financial Quarter
ics_table_formatted <- ics_prescribed %>% filter(fin_quarter %in% latest_full_quarter) %>%
  mutate(percent_highstrength = round(percent_highstrength,2)) %>%
  arrange(desc(percent_highstrength)) %>%
  rename(`Quantity - High Strength ICS` = highstrengthquantity,
         `Quantity - All ICS` = total_quantity,
         `Percentage of ICS that is High Strength` = percent_highstrength) %>%
  select(-fin_quarter)

# Summary of ICS high strength prescribing - overall
ics_totals <- ics_prescribed %>% group_by(fin_quarter) %>% 
  summarise(total_quantity = sum(total_quantity),
            highstrengthquantity = sum(highstrengthquantity),
            percent_highstrength = highstrengthquantity / total_quantity * 100)

ics_totals_longer <- ics_totals %>% rename(`All ICS` = total_quantity, `High Strength ICS` = highstrengthquantity) %>% 
  pivot_longer(cols = c(`All ICS`, `High Strength ICS`), names_to = "group", values_to = "quantity")

#------------ DPI vs MDI vs Soft Mist Inhaler Quantities over time---------------#

# Quantity and Items by Inhaler Type by Quarter
inhaler_types_by_quantity <- respiratory %>%
  filter(pi_drug_formulation == "INHAL", !fin_quarter %in% latest_incomplete_quarter) %>%
  left_join(inhaler_types, by = "pi_bnf_item_description") %>%
  group_by(fin_quarter, category) %>%
  summarise(quantity = sum(paid_quantity),
            items = sum(number_of_paid_items))

inhaler_types_by_percentage <- inhaler_types_by_quantity %>% select(-items) %>%
  pivot_wider(names_from = category, values_from = quantity) %>%
  mutate(total = DPI + MDI + `soft mist`,
         `Percentage of DPI Inhalers` = scales::percent(DPI / total, accuracy = .1),
         `Percentage of MDI Inhalers` = scales::percent(MDI / total, accuracy = .1),
         `Percentage of Soft Mist Inhalers` = scales::percent(`soft mist` / total, accuracy = .1)) %>%
  select(-c(total, DPI, MDI, `soft mist`)) %>%
  rename(`Financial Quarter` = fin_quarter)

# Extract list of inhalers to check
# inhaler_caps_list <- respiratory %>% filter(date >= full_back12Mth) %>%
#   filter(pi_drug_formulation == "CAPS") %>%
#   left_join(inhaler_types, by = "pi_bnf_item_description") %>%
#   group_by(inhaler_class, pi_approved_name, pi_prescribable_item_name, pi_bnf_item_description, pi_drug_formulation, category) %>% 
#   summarise(quantity = sum(paid_quantity),
#             items = sum(number_of_paid_items))

#-------------------------------- Clenil to Soprobec Switch -------------------------------#

# Summarise Soprobec and Clenil number of items from Jan 2020
soprobec_clenil_summary <- respiratory %>% filter(str_detect(pi_prescribable_item_name, "CLENIL") | str_detect(pi_prescribable_item_name, "SOPROBEC"), date >= "2020-01-01") %>%
  select(presc_location_name, pat_upi_c, date, pi_prescribable_item_name, pi_bnf_item_description, number_of_paid_items) %>% 
  # Classify items as Clenil or Soprobec
  mutate(class = case_when(
    str_detect(pi_prescribable_item_name, "CLENIL") ~ "Clenil",
    str_detect(pi_prescribable_item_name, "SOPROBEC") ~ "Soprobec"
  )) %>%
  group_by(date, class) %>%
  summarise(items = sum(number_of_paid_items)) %>%
  mutate(date = as.Date(date))

# Summarise Soprobec and Clenil items by Age
soprobec_clenil_summary_by_age <- respiratory %>% filter(str_detect(pi_prescribable_item_name, "CLENIL") | str_detect(pi_prescribable_item_name, "SOPROBEC"),
                                                         date >= "2020-01-01", !is.na(pat_date_of_birth_c)) %>%
  select(presc_location_name, pat_upi_c, pat_date_of_birth_c, date, pi_prescribable_item_name, pi_bnf_item_description, number_of_paid_items) %>%
  mutate(prescribed_age = as.period(interval(start = pat_date_of_birth_c, end = date))$year,
  class = case_when(
    str_detect(pi_prescribable_item_name, "CLENIL") ~ "Clenil",
    str_detect(pi_prescribable_item_name, "SOPROBEC") ~ "Soprobec"),
  age_group = case_when(
    prescribed_age < 18 ~ "Under 18",
    prescribed_age >= 18 ~ "Over 18"
  )) %>%
  group_by(date, class, age_group) %>%
  summarise(items = sum(number_of_paid_items))%>%
  mutate(date = as.Date(date))

#-------------------------------- Montelukast -------------------------------#

# Montelukast - Items, Quantity and Cost by Month
montelukast_summary <- montelukast %>% mutate(date = as.Date(paid_date, "%m/%d/%y")) %>%
  group_by(date) %>%
  summarise(items = sum(number_of_paid_items),
            quantity = sum(paid_quantity),
            cost = sum(pd_paid_gic_excl_bb),
            patients = n_distinct(pat_upi_c))
```

# Summary

## Row {data-height="100"}

### Total Respiratory patients in last 12 months, `r full_back12Mth_pretty` - `r latest_date`

```{r echo = F, message = F}
valueBox(format(total_respiratory_patients$patients, big.mark = ","), icon = "fa-users")
```


## Row
\ 
Author: Lara Johnson, Senior Information Analyst

Date: `r format(Sys.time(), '%d %B, %Y')` \

\
This report summarises inhaler prescribing trends in NHS Lothian.\
\
All prescribing data has been taken from the Prescribing Information System (PIS).\
\
[**Disclaimer: This file contains PIS data. The data presented have not been adjusted to protect against potential disclosure risks and are released for clinical and management information purposes. The data presented may contain information which enables (perhaps with the aid of further knowledge of the topic) an individual patient to be identified. Please ensure access is restricted and that patient confidentiality is not compromised.**]{style="color:red"}\

# 1 - Spend

Row {.tabset .tabset-fade} 
-------------------------------------
### 1.1 Spend by BNF Chapter
```{r fig 1.1, echo = F, message = F}
# Tree map of spend by BNF Chapter - last 12 months
treemap::treemap(bnf_spend, index = "label", vSize = "pd_paid_gic_excl_bb", type = "index", 
                 title = "Annual Spend by BNF Chapter")
```

### 1.2 Table - Spend on Inhalers in 12 Months
```{r fig 1.2, echo = F, message = F}
inhaler_spend %>% kbl() %>% kable_styling()
```

### 1.3 Bar Chart - Spend per Patient by Cluster
```{r fig 1.3, echo = F, message = F}
ggplot(respiratory_spend_bycluster)+
  geom_bar(aes(x = cluster, y = cost_per_treated_patient, fill = cluster), stat = "identity", show.legend = F)+
  geom_hline(aes(yintercept = mean(cost_per_treated_patient)))+
  geom_text(aes(1.5, mean(cost_per_treated_patient), label = "Lothian Mean", vjust = -1))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "Cost per Treated Patient",labels = scales::dollar_format(prefix = "£"))+
  labs(title = "Spend on inhalers by GP cluster", 
       caption = paste(format(full_back3Mth_pretty), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### 1.4 Scatterplot - Spend per Patient by Practice
```{r fig 1.4, echo = F, message = F}
ggplot(respiratory_bypractice)+
  geom_point(aes(x = respiratory_patients, y = cost_per_treated_patient, colour = cluster), show.legend = F)+
  geom_hline(aes(yintercept = mean(cost_per_treated_patient)))+
  geom_text(aes(110, mean(cost_per_treated_patient), label = "Lothian Mean", vjust = -1))+
  coord_cartesian(ylim = c(0,100))+
  theme_minimal()+
  scale_x_continuous(name = "Number of Treated Patients", breaks = seq(0,1500,100))+
  scale_y_continuous(name = "Cost per Treated Patient",labels = scales::dollar_format(prefix = "£"))+
  labs(title = "Spend on Inhalers - All Practices", 
       caption = paste(format(full_back3Mth_pretty), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### 1.5 Scatterplot - Items per Patient by Practice
```{r fig 1.5, echo = F, message = F}
# Items per treated patient
ggplot(respiratory_bypractice)+
  geom_point(aes(x = respiratory_patients, y = items_per_treated_patient, colour = cluster), show.legend = F)+
  geom_hline(aes(yintercept = mean(items_per_treated_patient)))+
  geom_text(aes(110, mean(items_per_treated_patient), label = "LothianMean", vjust = -1))+
  theme_minimal()+
  scale_x_continuous(name = "Number of Treated Patients", breaks = seq(0,1500,100))+
  scale_y_continuous(name = "Items per Treated Patient")+
  labs(title = "Number of Items - All Practices", 
       caption = paste(format(full_back3Mth_pretty), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```


# 2 - SABA Overuse

Row {.tabset .tabset-fade} 
-------------------------------------
### 2.1 Table - Single SABA Use
```{r fig 2.1, echo = F, message = F}
single_saba_use %>% kbl() %>% kable_styling()
```

### 2.2 Table - Single SABA Use by Practice
```{r fig 2.2, echo = F, message = F}
DT::datatable(single_saba_by_practice, rownames = F, filter = 'top', class = 'cell-border stripe', editable = T)
```

### 2.3 Box Plot (> 12 SABA)
```{r fig 2.3, echo = F, message = F}
# Box Plot - % Patients Prescribed > 12 SABA per annum as a distribution by locality
ggplot(SABAoveruse, aes(x = locality, y = percent12))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(alpha = .6, width = .15, aes(colour = locality), show.legend = F)+
  geom_hline(aes(yintercept = mean(percent12), linetype = "Lothian Mean"), colour = "grey", size = 1.2)+
  theme_minimal()+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "% Patients with > 12 SABA in 12 Months")+
  labs(title = "Percentage of Patients Prescribed > 12 SABA per annum", 
       caption = paste(format(full_back12Mth_pretty), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### 2.4 Box Plot (>6 SABA)
```{r fig 2.4, echo = F, message = F}
# Percentage of patients prescribed > 6 SABA per annum
ggplot(SABAoveruse, aes(x = locality, y = percent6))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(alpha = .6, width = .15, aes(colour = locality), show.legend = F)+
  geom_hline(aes(yintercept = mean(percent6), linetype = "Lothian Mean"), colour = "grey", size = 1.2)+
  theme_minimal()+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "% Patients with > 6 SABA in 12 Months")+
  labs(title = "Percentage of Patients Prescribed > 6 SABA per annum", 
       caption = paste(format(full_back12Mth_pretty), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### 2.5 Table - SABA Overuse by Practice
```{r fig 2.5, echo = F, message = F}
# Table of SABA overuse by practices
DT::datatable(SABAoveruse_formatted, rownames = F, filter = 'top', class = 'cell-border stripe', editable = T)
```

### 2.6 Run Chart (>12 SABA)
```{r fig 2.6, echo = F, message = F}
# Run chart of SABA overuse over time
fig2_6 <- ggplot(SABAoveruse_by_year, aes(x = fin_year, y = percent12))+
  geom_line(aes(group = presc_location_name, colour = locality))+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "% Patients with > 12 SABA in 12 Months")+
  theme(axis.line = element_line(colour = "black"))+
  facet_wrap(~locality)+
  labs(title = "% Patients with > 12 SABA in 12 Months in NHS Lothian", 
       caption = paste("January 2018", "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))

ggplotly(fig2_6)
```

### 2.7 Run Chart (>6 SABA)
```{r fig 2.7, echo = F, message = F}
# Run chart of SABA overuse over time
fig2_7 <- ggplot(SABAoveruse_by_year, aes(x = fin_year, y = percent6))+
  geom_line(aes(group = presc_location_name, colour = locality))+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "% Patients with > 6 SABA in 12 Months")+
  theme(axis.line = element_line(colour = "black"))+
  facet_wrap(~locality)+
  labs(title = "% Patients with > 6 SABA in 12 Months in NHS Lothian", 
       caption = paste("January 2018", "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))

ggplotly(fig2_7)
```

### 2.8 More than 2 SABA MDI per Prescription
```{r fig 2.8, echo = F, message = F}
# Number of Patients on > 200 doses of SABA MDI (200 doses = 1 inhaler), broken down by practice
DT::datatable(SABAdoses, rownames = F, filter = 'top', class = 'cell-border stripe', editable = T, extensions = 'Buttons', 
              options = list(dom = 'Blfrtip', buttons = c('copy', 'excel')))

```




# 3 - High Strength ICS

Row {.tabset .tabset-fade} 
-------------------------------------
### 3.1 High Strength ICS - Last 3 Months
```{r fig 3.1, echo = F, message = F}
# Box plot of ICS percentage in most recent full quarter
ggplot(ics_prescribed[ics_prescribed$fin_quarter %in% latest_full_quarter,], aes(x = locality, y = percent_highstrength))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter(alpha = .6, width = .15, aes(colour = locality), show.legend = F)+
  geom_hline(aes(yintercept = mean(percent_highstrength), linetype = "Lothian Mean"), colour = "grey", size = 1.2)+
  theme_minimal()+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "% ICS Prescribed as High Strength per Practice")+
  labs(title = "Percentage of ICS Classified as High Strength", 
       caption = paste(format(full_back3Mth_pretty), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### 3.2 Table - High Strength ICS by practice
```{r fig 3.2, echo = F, message = F}
DT::datatable(ics_table_formatted, rownames = F, filter = 'top', class = 'cell-border stripe', editable = T)
```

### 3.3 High Strength ICS Run Chart
```{r fig 3.3, echo = F, message = F}
# Time Series Trends
# Fig 3.3 with bar chart on secondary axis
ggplot(ics_totals_longer, aes(x = fin_quarter))+
  geom_line(aes(y = percent_highstrength, group = 1))+
  geom_col(aes(y = quantity / 10000, fill = group), position = "dodge")+
  scale_y_continuous(name = "% ICS Prescribed as High Strength", labels = scales::percent_format(scale = 1),
                     # Format secondary axis
                     sec.axis = sec_axis(~. * 10000, name = "Quantity", labels = scales::comma))+
  theme_minimal()+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  labs(title = "% of ICS Classified as High Strength in NHS Lothian", 
     caption = paste("January 2018", "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```


# 4 - DPI & MDI Inhalers

Row {.tabset .tabset-fade} 
-------------------------------------

### Fig 4.1 Quantity
```{r fig 4.1, echo = F, message = F}
# Quantity
ggplot(inhaler_types_by_quantity)+
  geom_line(aes(x = fin_quarter, y = quantity, group = category, colour = category))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "Quantity of Inhalers", labels = scales::comma)+
  labs(title = "Quantity of DPI, MDI and Soft Mist Inhalers\n Prescribed Over Time", 
       caption = paste("April 2016", "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### Fig 4.2 Items
```{r fig 4.2, echo = F, message = F}
# Number of Items
ggplot(inhaler_types_by_quantity)+
  geom_line(aes(x = fin_quarter, y = items, group = category, colour = category))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "Number of Items", labels = scales::comma)+
  labs(title = "Number of Items of DPI, MDI and Soft Mist Inhalers\n Prescribed Over Time", 
       caption = paste("April 2016", "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))
```

### Fig 4.3 Percentages
```{r fig 4.3, echo = F, message = F}
DT::datatable(inhaler_types_by_percentage, rownames = F, filter = 'top', class = 'cell-border stripe', editable = T)
```


# 5 - Clenil - Soprabec

Row {.tabset .tabset-fade} 
-------------------------------------

### 5.1 Run Chart
```{r fig 5.1, echo = F, message = F}
# Run chart for Soprobec and Clenil number of items from Jan 2020
ggplot(soprobec_clenil_summary)+
  geom_line(aes(x = date, y = items, group = class, colour = class), size = 1)+
  geom_vline(aes(xintercept = as.Date("2020-03-01")), linetype = "dashed", size = 1)+
  annotate(geom = "text", x = as.Date("2020-01-22"), y = max(soprobec_clenil_summary$items), label = "Switch Start\n(Mar 2020)")+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())+
  scale_y_continuous(name = "Number of Prescriptions", labels = scales::comma)+
  scale_x_date(name = " ")+
  theme(legend.title = element_blank())+
  labs(title = "Number of Clenil and Soprobec Items Prescribed", 
       caption = paste(format(min(soprobec_clenil_summary$date, na.rm = T), "%B %Y"), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))

```

### 5.2 Run Chart by Age
```{r fig 5.2, echo = F, message = F}
# Run chart for Soprobec and Clenil - Under vs Over 18s
ggplot(soprobec_clenil_summary_by_age)+
  geom_line(aes(x = date, y = items, group = class, colour = class), size = 1)+
  geom_vline(aes(xintercept = as.Date("2020-03-01")), linetype = "dashed", size = 1)+
  annotate(geom = "text", x = as.Date("2020-01-22"), y = max(soprobec_clenil_summary$items) - 4000, label = "Switch Start\n(Mar 2020)")+
  theme_minimal()+
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())+
  scale_y_continuous(name = "Number of Prescriptions", labels = scales::comma)+
  scale_x_date(name = " ")+
  #theme(legend.title = element_blank())+
  theme(legend.title = element_blank(), legend.position = "bottom")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~age_group, nrow = 2)+
  labs(title = "Number of Clenil and Soprobec Items Prescribed by Age", 
       caption = paste(format(min(soprobec_clenil_summary$date, na.rm = T), "%B %Y"), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))

```


# 6 - Montelukast

Row {.tabset .tabset-fade} 
-------------------------------------
### 6.1 Quantity
```{r fig 6.1, echo = F, message = F}
# Montelukast - run chart of prescribing quantity of all strengths of montelukast
ggplot(montelukast_summary, aes(x = date))+
  geom_line(aes(y = quantity), size = 1)+
  geom_col(aes(y = patients * 30))+
  scale_y_continuous(name = "Paid Quantity of Montelukast", labels = scales::comma,
                     #Format secondary axis
                     sec.axis = sec_axis(~. / 30, name = "Number of Patients", breaks = seq(0,3000, 500)))+
  theme_minimal()+
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months", name = " ")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  labs(title = "Quantity of Montelukast Prescribed Over Time", 
       caption = paste(format(min(montelukast_summary$date, na.rm = T), "%B %Y"), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))

```

### 6.2 Spend on Montelukast
```{r fig 6.2, echo = F, message = F}
# Montelukast - run chart of cost
ggplot(montelukast_summary, aes(x = date))+
  geom_line(aes(y = cost), size = 1)+
  geom_col(aes(y = patients * 3))+
  scale_y_continuous(name = "Paid Cost of Montelukast (£)", labels = scales::comma,
                     #Format secondary axis
                     sec.axis = sec_axis(~. / 3, name = "Number of Patients", breaks = seq(0,3000, 500)))+
  theme_minimal()+
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months", name = " ")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  labs(title = "Spend on Montelukast Prescribed Over Time", 
       caption = paste(format(min(montelukast_summary$date, na.rm = T), "%B %Y"), "-", format(latest_date), sep = " ", "| NHS Lothian | Data Source: PIS"))

```





