---
title: "Prescribing Indicators"
author: "Lara"
date: "19/01/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
runtime: shiny
---

```{r setup, echo =FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(readr)
library(dplyr)
library(dbplyr)
library(lubridate)
library(readxl)
library(ggplot2)
library(plotly)
library(shiny)
library(magrittr)
library(DT)
```
# Antibiotics
```{r antibiotics, echo = F}
#=============================================
# Process the Antibiotics Dataset (for Indicators 1-3)
#=============================================

# Load in antibiotics dataset
antibiotics <- fread("H:\\Training datasets\\Prescribing_indicators\\1-Antibiotics.csv")
colnames(antibiotics) <- snakecase::to_snake_case(colnames(antibiotics)) #Rename columns
# Change some strings to numeric and dates (be careful of century when converting 2-digit years to 4-digit years)
antibiotics %<>% mutate(pat_date_of_birth_c = mdy(pat_date_of_birth_c),
                        pat_date_of_birth_c = ifelse(year(pat_date_of_birth_c) > 2021,
                                                     paste(subtract((year(pat_date_of_birth_c)), 100),
                                                           format(pat_date_of_birth_c, "%m-%d"),
                                                           sep = "-"),
                                                     as.character(pat_date_of_birth_c)),
                        pat_date_of_birth_c = as.POSIXct(pat_date_of_birth_c),
                        paid_date = as.POSIXct(paid_date, format = "%m/%d/%y"),
                        paid_quantity = as.numeric(gsub(",", "", paid_quantity)),
                        pd_paid_gic_excl_bb = as.numeric(gsub(",", "", pd_paid_gic_excl_bb)),
                        paid_month_year = zoo::as.yearmon(paid_date, "%b %Y"),
                        paid_financial_quarter = case_when(
                          paid_date >= "2019-10-01" & paid_date <= "2019-12-31" ~ "Q3 19/20",
                          paid_date >= "2020-01-01" & paid_date <= "2020-03-31" ~ "Q4 19/20",
                          paid_date >= "2020-04-01" & paid_date <= "2020-06-30" ~ "Q1 20/21",
                          paid_date >= "2020-07-01" & paid_date <= "2020-09-30" ~ "Q2 20/21",
                          paid_date >= "2020-10-01" & paid_date <= "2020-12-31" ~ "Q3 20/21",
                          paid_date >= "2021-01-01" & paid_date <= "2021-03-31" ~ "Q4 20/21")
                        )

#Load a list of active GPs during the reporting period and map their localities / clusters
source("N:/Projects/Primary Care Pharmacy/Tips and Training/R/loadPharmacyFunctions.R")

practices <- antibiotics %>% select(presc_location_code) %>% distinct()
practiceLookup <- searchLookupFile(practiceCode = practices$presc_location_code) %>%
  select(PracticeCode, locality, cluster)

# Add practice population (as of Jan 2020)
practice_population <- read_excel("H:\\Training datasets\\Practice_pop_2021-01-15.xlsx")
practice_population %<>% select(LOCATION_CODE, PRACTICE_POPULATION)
colnames(practice_population) <- snakecase::to_snake_case(colnames(practice_population))

practiceLookup %<>% left_join(practice_population, by = c("PracticeCode" = "location_code"))

# Add column of real practice list size and population by age bands TO DO
# Connect to TRAK and run SQL file


#Use the gpLookup file to match on locality/cluster
antibiotics %<>% left_join(practiceLookup,
                              by = c("presc_location_code" = "PracticeCode"))

# Remove NAs from locality and practice_population
antibiotics %<>% filter(!is.na(locality), !is.na(practice_population), locality != "No HSCP Mapping Found")
```

# Indicator 1 - Antibiotic Items per 1000 list size per day ≤ 1.64
```{r indicator 1, echo = F}
#knitr::opts_chunk$set(echo = TRUE)
# Indicator 1 - summarise by financial quarter
indicator1_byquarter <- antibiotics %>% group_by(locality, cluster, presc_location_code, presc_location_name, practice_population, paid_financial_quarter) %>%
  summarise(antibiotics_items = sum(number_of_paid_items)) %>%
  mutate(items_per_1000_listsize = round((antibiotics_items*1000)/practice_population,2),
         items_per_1000_listsize_perday = round(items_per_1000_listsize/(365/4),2),
         target = 1.64,
         target_met = if_else(items_per_1000_listsize_perday <= target, "Yes", "No"))

# Indicator 1 - summarise by month
indicator1_bymonth <- antibiotics %>% group_by(locality, cluster, presc_location_code, presc_location_name, practice_population, paid_financial_quarter, paid_month_year, paid_date) %>%
  summarise(antibiotics_items = sum(number_of_paid_items)) %>%
  mutate(items_per_1000_listsize = round((antibiotics_items*1000)/practice_population,2),
         items_per_1000_listsize_perday = round(items_per_1000_listsize/(365/12),2),
         target = 1.64,
         target_met = if_else(items_per_1000_listsize_perday <= target, "Yes", "No"))

# Extract baseline figures
ind1_baseline <- indicator1_byquarter %>% filter(paid_financial_quarter == "Q3 19/20") %>% 
  select(presc_location_code, presc_location_name, items_per_1000_listsize_perday)
ind1_baseline$baseline <- ind1_baseline$items_per_1000_listsize_perday
ind1_baseline <- ind1_baseline[,c(4,7)]

# Locality means by month
ind1_localitymeans <- indicator1_bymonth %>% group_by(locality, paid_date, paid_month_year, target) %>%
  summarise_at(vars(items_per_1000_listsize_perday), list(localitymean = mean))  %>%
  mutate(localitymean = round(localitymean, 2))

# Locality means by quarter
ind1_localitymeans_byquarter <- indicator1_byquarter %>% group_by(locality, paid_financial_quarter, target) %>%
  summarise_at(vars(items_per_1000_listsize_perday), list(localitymean = mean))  %>%
  mutate(localitymean = round(localitymean, 2))

```

## Table
```{r echo = F}
inputPanel(
  selectInput("quarter3", label = "Select a quarter:",
              choices = c(indicator1_byquarter$paid_financial_quarter %>% unique())
                ))
  DT::renderDataTable({
    indicator1_byquarter %>% filter(paid_financial_quarter == input$quarter3) %>%
      select(!c(paid_financial_quarter, items_per_1000_listsize, target)) %>%
      left_join(ind1_baseline, by = "presc_location_code") %>%
      # Add columns for practice baseline and % change since baseline
  mutate(baseline_change = items_per_1000_listsize_perday - baseline,
         percentage_change = round(baseline_change / baseline,2)) 
    }, rownames = F, filter = "top",
  colnames = c("Locality", "Cluster", "Location Code", "Location Name", "Practice Population", "Items Prescribed", "Items per 1000 List Size Per Day", "Target Met", "Baseline", "Change from Baseline", "Percentage Reduction"),
    extensions = 'Buttons', options = list(dom = 'Bfrtrip', buttons = c('copy', 'excel','csv', 'pdf'))
)
```

## Bar Chart
### Antibiotics Items Prescribed by Practice by Quarter 
```{r table, echo = F}
inputPanel(
# Selector for choosing locality
  selectInput("locality", label = "Choose a locality:",
              choices = c("East Lothian", "Midlothian", "North East Edinburgh", "North West Edinburgh", "South East Edinburgh", "South West Edinburgh", "West Lothian")),
# Selector for choosing quarter
  selectInput("quarter", label = "Choose a quarter:",
              choices = c(indicator1_byquarter$paid_financial_quarter %>% unique())
))

renderPlot({
  ggplot(indicator1_byquarter[indicator1_byquarter$locality == input$locality & indicator1_byquarter$paid_financial_quarter == input$quarter,])+
  geom_bar(aes(x = reorder(presc_location_name, items_per_1000_listsize_perday), y = items_per_1000_listsize_perday, fill = target_met), stat = "identity")+
  geom_hline(aes(yintercept = target, linetype = "Target (1.64)"))+
  geom_hline(aes(yintercept = mean(items_per_1000_listsize_perday), linetype = "Locality Mean"))+
  scale_fill_discrete(name = "Target Met?", breaks = c("No", "Yes"))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "Antibiotics Items Prescribed Per 1000 List Size")+
  theme_minimal()+
  coord_flip()
})

```

## Trends Over Time
### Means per Locality by Quarter
```{r timeseries, echo = F}
# Time Series - Overview of All Localities
ggplot(ind1_localitymeans)+
  geom_line(aes(x = paid_date, y = localitymean, colour = locality))+
  geom_hline(aes(yintercept = target))+
  annotate("text", min(ind1_localitymeans$paid_date), 1.64, vjust = -.5, label = "Target")+
  scale_y_continuous(name = "Locality Mean")+
  theme_minimal()
```

### Table of Locality Means by Quarter
```{r echo = F}
inputPanel(
  selectInput("ind1_quarter2", label = "Select a quarter:",
              choices = c(ind1_localitymeans_byquarter$paid_financial_quarter %>% unique())
                ))
  DT::renderDataTable({
    ind1_localitymeans_byquarter[ind1_localitymeans_byquarter$paid_financial_quarter == input$ind1_quarter2,]}, rownames = F, filter = "top", 
             colnames = c("Locality", "Paid Financial Quarter", "Target", "Locality Mean"),
    extensions = 'Buttons', options = list(dom = 'Bfrtrip', buttons = c('copy', 'excel','csv', 'pdf'))
)

```

### All Practices Within a Locality Over Time
```{r timeseries_locality, echo = F}
inputPanel(
# Selector for choosing locality
  selectInput("locality2", label = "Choose a locality:",
              choices = c("East Lothian", "Midlothian", "North East Edinburgh", "North West Edinburgh", "South East Edinburgh", "South West Edinburgh", "West Lothian"))
)

renderPlot({
  ggplot(indicator1_bymonth[indicator1_bymonth$locality == input$locality2,])+
  geom_line(aes(x = paid_date, y = items_per_1000_listsize_perday, group = presc_location_name, colour = presc_location_name), alpha = .7)+
  geom_hline(aes(yintercept = target))+
  annotate("text", min(ind1_localitymeans$paid_date), 1.64, vjust = -.5, label = "Target")+
  geom_line(data = ind1_localitymeans[ind1_localitymeans$locality == input$locality2,], aes(x = paid_date, y = localitymean), colour = "darkred", size = 2)+
  scale_y_continuous(name = "Antibiotics Items Prescribed Per 1000 List Size")+
  theme_minimal()
})
```

### Individual Practices Over Time
```{r timeseries_practices, echo = F}
inputPanel(
# Selector for choosing locality
  selectInput("locality3", label = "Choose a Practice:",
              choices = c(indicator1_bymonth$presc_location_name %>% sort() %>% unique())
))

renderPlot({
  ggplot(indicator1_bymonth[indicator1_bymonth$presc_location_name == input$locality3,])+
  geom_line(aes(x = paid_date, y = items_per_1000_listsize_perday, colour = presc_location_name))+
  geom_hline(aes(yintercept = target))+
  annotate("text", min(ind1_localitymeans$paid_date), 1.64, vjust = -.5, label = "Target")+
  geom_line(data = ind1_localitymeans[ind1_localitymeans$locality == indicator1_bymonth$locality[indicator1_bymonth$presc_location_name == input$locality3],], aes(x = paid_date, y = localitymean, colour = "Locality Mean"))+
  scale_y_continuous(name = "Antibiotics Items Prescribed Per 1000 List Size")+
  theme_minimal()
})
  
```

# Indicator 2 - Number of Co-Amoxiclav Items per 1000 List Size per 100 Days ≤ 6.00
```{r indicator2, echo = F}
co_amoxiclav <- antibiotics %>% filter(pi_approved_name == "CO-AMOXICLAV")

# Indicator 2 - summarise by financial quarter
indicator2_byquarter <- co_amoxiclav %>% group_by(locality, cluster, presc_location_code, presc_location_name, practice_population, paid_financial_quarter) %>%
  summarise(antibiotics_items = sum(number_of_paid_items)) %>%
  mutate(items_per_1000_listsize = round((antibiotics_items*1000)/practice_population,2),
         items_per_1000_listsize_per100days = round((items_per_1000_listsize/(365/4)*100),2),
         target = 6.00,
         target_met = if_else(items_per_1000_listsize_per100days <= target, "Yes", "No"))

# Indicator 1 - summarise by month
indicator2_bymonth <- co_amoxiclav %>% group_by(locality, cluster, presc_location_code, presc_location_name, practice_population, paid_financial_quarter, paid_month_year, paid_date) %>%
  summarise(antibiotics_items = sum(number_of_paid_items)) %>%
  mutate(items_per_1000_listsize = round((antibiotics_items*1000)/practice_population,2),
         items_per_1000_listsize_per100days = round((items_per_1000_listsize/(365/12)*100),2),
         target = 6.00,
         target_met = if_else(items_per_1000_listsize_per100days <= target, "Yes", "No"))

# Extract baseline figures
ind2_baseline <- indicator2_byquarter %>% filter(paid_financial_quarter == "Q3 19/20") %>% 
  select(presc_location_code, presc_location_name, items_per_1000_listsize_per100days)
ind2_baseline$baseline <- round(ind2_baseline$items_per_1000_listsize_per100days,2)
ind2_baseline <- ind2_baseline[,c(4,7)]

# Locality means by month
ind2_localitymeans <- indicator2_bymonth %>% group_by(locality, paid_date, paid_month_year, target) %>%
  summarise_at(vars(items_per_1000_listsize_per100days), list(localitymean = mean))  %>%
  mutate(localitymean = round(localitymean, 2))

# Locality means by quarter
ind2_localitymeans_byquarter <- indicator2_byquarter %>% group_by(locality, paid_financial_quarter, target) %>%
  summarise_at(vars(items_per_1000_listsize_per100days), list(localitymean = mean))  %>%
  mutate(localitymean = round(localitymean, 2))
```

## Table
```{r echo = F}
inputPanel(
  selectInput("ind2_quarter", label = "Select a quarter:",
              choices = c(indicator2_byquarter$paid_financial_quarter %>% unique())
                ))
  DT::renderDataTable({
    indicator2_byquarter %>% filter(paid_financial_quarter == input$ind2_quarter) %>%
      select(!c(paid_financial_quarter, items_per_1000_listsize, target)) %>%
      left_join(ind2_baseline, by = "presc_location_code") %>%
      # Add columns for practice baseline and % change since baseline
  mutate(baseline_change = items_per_1000_listsize_per100days - baseline,
         percentage_change = round(baseline_change / baseline,2)) 
    }, rownames = F, filter = "top",
  colnames = c("Locality", "Cluster", "Location Code", "Location Name", "Practice Population", "Items Prescribed", "Items per 1000 List Size Per 100 Days", "Target Met", "Baseline", "Change from Baseline", "Percentage Reduction"),
    extensions = 'Buttons', options = list(dom = 'Bfrtrip', buttons = c('copy', 'excel','csv', 'pdf'))
)
```

## Bar Chart
### Co-Amoxiclav Items Prescribed by Practice by Quarter 
```{r echo = F}
inputPanel(
# Selector for choosing locality
  selectInput("ind2_locality", label = "Choose a locality:",
              choices = c("East Lothian", "Midlothian", "North East Edinburgh", "North West Edinburgh", "South East Edinburgh", "South West Edinburgh", "West Lothian")),
# Selector for choosing quarter
  selectInput("ind2_quarter1", label = "Choose a quarter:",
              choices = c(indicator2_byquarter$paid_financial_quarter %>% unique())
))

renderPlot({
  ggplot(indicator2_byquarter[indicator2_byquarter$locality == input$ind2_locality & indicator2_byquarter$paid_financial_quarter == input$ind2_quarter1,])+
  geom_bar(aes(x = reorder(presc_location_name, items_per_1000_listsize_per100days), y = items_per_1000_listsize_per100days, fill = target_met), stat = "identity")+
  geom_hline(aes(yintercept = target, linetype = "Target (6.00)"))+
  geom_hline(aes(yintercept = mean(items_per_1000_listsize_per100days), linetype = "Locality Mean"))+
  scale_fill_discrete(name = "Target Met?", breaks = c("No", "Yes"))+
  scale_x_discrete(name = "")+
  scale_y_continuous(name = "Co-Amoxiclav Items Prescribed Per 1000 List Size")+
  theme_minimal()+
  coord_flip()
})
```


## Trends Over Time
### Means per Locality by Quarter
```{r echo = F}
# Time Series - Overview of All Localities
ggplot(ind2_localitymeans)+
  geom_line(aes(x = paid_date, y = localitymean, colour = locality))+
  geom_hline(aes(yintercept = target))+
  annotate("text", min(ind2_localitymeans$paid_date), 6.00, vjust = -.5, label = "Target")+
  scale_y_continuous(name = "Locality Mean")+
  theme_minimal()
```

### Table of Locality Means by Quarter
```{r echo = F}
inputPanel(
  selectInput("ind2_quarter2", label = "Select a quarter:",
              choices = c(ind2_localitymeans_byquarter$paid_financial_quarter %>% unique())
                ))
  DT::renderDataTable({
    ind2_localitymeans_byquarter[ind2_localitymeans_byquarter$paid_financial_quarter == input$ind2_quarter2,]}, rownames = F, filter = "top", 
             colnames = c("Locality", "Paid Financial Quarter", "Target", "Locality Mean"),
    extensions = 'Buttons', options = list(dom = 'Bfrtrip', buttons = c('copy', 'excel','csv', 'pdf'))
)

```

### All Practices Within a Locality Over Time
```{r echo = F}
inputPanel(
# Selector for choosing locality
  selectInput("ind2_locality2", label = "Choose a locality:",
              choices = c("East Lothian", "Midlothian", "North East Edinburgh", "North West Edinburgh", "South East Edinburgh", "South West Edinburgh", "West Lothian"))
)

renderPlot({
  ggplot(indicator2_bymonth[indicator2_bymonth$locality == input$ind2_locality2,])+
  geom_line(aes(x = paid_date, y = items_per_1000_listsize_per100days, group = presc_location_name, colour = presc_location_name), alpha = .7)+
  geom_hline(aes(yintercept = target))+
  annotate("text", min(ind2_localitymeans$paid_date), 1.64, vjust = -.5, label = "Target")+
  geom_line(data = ind2_localitymeans[ind2_localitymeans$locality == input$ind2_locality2,], aes(x = paid_date, y = localitymean), colour = "darkred", size = 2)+
  scale_y_continuous(name = "Co-Amoxiclav Items Prescribed Per 1000 List Size")+
  theme_minimal()
})
```

### Individual Practices Over Time
```{r echo = F}
inputPanel(
# Selector for choosing locality
  selectInput("ind2_locality3", label = "Choose a Practice:",
              choices = c(indicator1_bymonth$presc_location_name %>% sort() %>% unique())
))

renderPlot({
  ggplot(indicator2_bymonth[indicator2_bymonth$presc_location_name == input$ind2_locality3,])+
  geom_line(aes(x = paid_date, y = items_per_1000_listsize_per100days, colour = presc_location_name))+
  geom_hline(aes(yintercept = target))+
  annotate("text", min(ind2_localitymeans$paid_date), 1.64, vjust = -.5, label = "Target")+
  geom_line(data = ind2_localitymeans[ind2_localitymeans$locality == indicator2_bymonth$locality[indicator2_bymonth$presc_location_name == input$ind2_locality3],], aes(x = paid_date, y = localitymean, colour = "Locality Mean"))+
  scale_y_continuous(name = "Co-Amoxiclav Items Prescribed Per 1000 List Size")+
  theme_minimal()
})
  
```



This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).



